Dark matter halo shapes in the Auriga simulations

### 2019-08-21 Document/paper.tex
edit

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_DM.pdf}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_MHD.pdf}
\end{center}
\caption{Axes ratios for all simulated halos.
  Left/right panel corresponds to DMO/MHD simulations, respectively.
  Triangles (squares) represent the measurements at $R_{200}/16$
  ($R_{200}$) that correspond to physical distances of $\approx14$ kpc
  ($\approx 230$ kpc).
  Here we visualize three main population trends.
  First, in DMO simulations halos are rounder in the outskirts
  than in the inner part.
  Second, halos in MHD are rounder in the inner regions than in
  the outskirts (opposite to the DMO trend). 
  Third, halos in the MHD simulations are rounder than their DMO counterparts.
  The first and second points are further 
  clarified in Figure~\ref{fig:triaxial_cumulative} through cumulative triaxiality
  distributions.}
  \label{fig:triaxiality_plane}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{Angles between the principal axis of the dark matter halo
  shape and the angular momentum of the stellar disc 
  as a function of the radius at
  which the halo shape directions are measured.
  Each panel compares the alignment of the corresponding
  major, middle, and minor axis in the halo.
  Thin lines correspond to each one of the thirty halos in the sample
  while the thick line traces the median value as a function of  radius.
  In the upper row, the haloes come from the DMO simulation.
  The first thing to notice is that the major axis is statistically
  perpendicular to the angular momentum of the disc that will form in
  the MHD simulations. The other two directions do not show any
  correlation. 
  The second important result is that the directions in the ellipsoids
  describing the shape are constant as a function of radius except for two
  cases. 
  In the lower row the haloes come from the MHD simulation, providing a 
  self-consistent comparison with the stellar discs. 
  In this case, the sample presents a good alignment of the angular
  momentum with the minor halo axis. 
  Furthermore, the dark matter shells twist significantly in $1/5$ of
  the sample. For the remaining haloes, the alignment is almost
  constant as a function of radius. 
}
\label{fig:cumulative_alignment}
\end{figure*}

@@

The left panel of Figure \ref{fig:triaxial_cumulative} shows that in
the DMO sample the median triaxiality is larger than $2/3$ (a
typical value that marks the transition from triaxiality to
prolateness) at all radii. Furthermore this median value increases 
as we move towards the inner part of the halo.
The middle panel in the same Figure~\ref{fig:triaxial_cumulative}
shows how the trends in the MHD simulations go 
in the opposite direction with respect to the DMO results.
There the median triaxiality is always smaller than $1/3$ (a typical
value markin the transition from triaxiality to prolateness) and this
median value decreases as we move towards the inner part of the halo.

@@

A common assumption in dynamical models of the MW dark matter halo is that
its angular momentum is perfectly aligned with the minor axis of the stellar disc.
Although this is a reasonable assumption due to the stability of
the galactic disc in this configuration in simplified models of isolated galaxies, this
might not perfectly hold in a full cosmological context. 
To examine the degree of validity of this assumption we study in this
subsection the alignment between the stellar disc's angular momentum
eigenvectors and the dark matter halo shape.

The upper row in Figure \ref{fig:cumulative_alignment} shows 
the alignments between the major, intermediate, and minor axis of the
halo and the disc angular momentum, as a function of the dimensionless radius at which the halo shape was measured. The upper row shows the angle between the DMO halo and the stellar
disk in the MHD simulation.
This comparison help us to measure how the shape directions in the
dark matter shells change with radius.
For the bulk of the sample the lines are horizontal.
Only two halos show a significant change of more than 30 degrees across different radii.

The lower row in Figure \ref{fig:cumulative_alignment} shows the
results for the MHD simulations.
The most striking feature of this plot is that most of the discs have their angular momentum aligned with the minor axis of the halo, perpendicular to the intermediate and major axis.
The median and the standard deviation of the alignment angles with
the major/median/minor axis are $82\pm14$, $82\pm16$ and $19\pm20$
degrees, respectively.

>>>>

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_DM.pdf}
\includegraphics[width=0.8\columnwidth]{Lvl_4_Triax_Plane_MHD.pdf}
\end{center}
\caption{Axes ratios for all simulated halos.
  Left/right panel corresponds to DMO/MHD simulations, respectively.
  Triangles (squares) represent the measurements at $R_{200}/16$
  ($R_{200}$) that correspond to physical distances of $\approx14$ kpc
  ($\approx 230$ kpc).
  Here we visualize three main population trends.
  First, in DMO simulations halos are rounder in the outskirts
  than in the inner part.
  Second, halos in MHD are rounder in the inner regions than in
  the outskirts (opposite to the DMO trend). 
  Third, halos in the MHD simulations are rounder than their DMO counterparts.
  These three points are further clarified in
  Figure~\ref{fig:triaxial_cumulative} through cumulative triaxiality distributions.}
  \label{fig:triaxiality_plane}
\end{figure*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{angles_alignment_MHD.pdf}
\end{center}
\caption{Angles between the principal axis of the dark matter halo
  shape and the angular momentum of the stellar disc 
  as a function of the radius at
  which the halo shape directions are measured.
  Each panel compares the alignment of the corresponding
  major, middle, and minor axis in the halo.
  Thin lines correspond to each one of the thirty halos in the sample
  while the thick line traces the median value as a function of  radius.
  The sample presents a good alignment of the angular
  momentum with the minor halo axis. 
  This alignment is also constant in radius.
  However, the dark matter shells twist significantly in $6$ halos 
  of our sample making the alignment measurement change. 
}
\label{fig:cumulative_alignment}
\end{figure*}

@@

The left panel of Figure \ref{fig:triaxial_cumulative} shows that in
the DMO sample the median triaxiality is larger than $2/3$ (a
typical value that marks the transition from triaxiality to
prolateness) at all radii. Furthermore this median value increases 
as we move towards the inner part of the halo.
The middle panel in the same Figure~\ref{fig:triaxial_cumulative}
shows how the trends in the MHD simulations go 
in the opposite direction with respect to the DMO results.
There the median triaxiality is always smaller than $1/3$ (a typical
value marking the transition from triaxiality to prolateness) and this
median value decreases as we move towards the inner part of the halo.

@@

A common assumption in dynamical models of the MW dark matter halo is that
its minor axis is perfectly aligned with the angular momentym of the
stellar disc. 
Although this is a reasonable assumption due to the stability of
the galactic disc in this configuration in simplified models of isolated galaxies, this
might not perfectly hold in a full cosmological context. 
To examine the degree of validity of this assumption we study in this
subsection the alignment between the stellar disc's angular momentum
eigenvectors and the dark matter halo shape.

Figure \ref{fig:cumulative_alignment} shows the
results as a function of radius.
The most striking feature of this plot is that most of the discs have
their angular momentum aligned with the minor axis of the halo,
perpendicular to the intermediate and major axis. 
The median and the standard deviation of the alignment angles with
the major/median/minor axis are $82\pm14$, $82\pm16$ and $19\pm20$
degrees, respectively.
This confirms statistically the expectation of a galactic disc aligned
with the minor axis of its dark matter halo, although some outliers
are present in the sample.
